\mytitle{}

### 2019-05-13 paper/paper.tex
Draft 1 for TNSM

\begin{abstract}
?? Trim to 75--200 words, simplify.

Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.
The detection problem in this domain presents unique difficulties; it is an evolving, non-stationary problem as usage patterns shift and new protocols and applications are introduced, further compounded by burstiness and seasonal variation.

\emph{Reinforcement learning} (RL) may overcome the detection problem for certain classes of anomaly by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available).
We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
We advance the state-of-the-art of DDoS mitigation using RL by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner.
This is supported by an in-depth investigation of feature suitability and empirical evaluation.

Our results show the existence of flow features with high predictive power for environments with predominantly TCP or UDP traffic, when such features are used as a basis for feedback-loop-like control.
We show that the new RL agent models can offer a significant increase in goodput of legitimate UDP traffic for many choices of host density, alongside a marked increase in protected TCP traffic at high host density.

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art, and we show that our second model achieves excellent performance for high host density.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Marl++}, \emph{SPF}), improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}) and discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

@@

\section{DDoS Mitigation with Per-flow Reinforcement Learning}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

@@

	\caption{
		Online performance for Opus benign traffic in a single-destination network, multi-agent mode.
%		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
		\emph{Guarded} always underperforms compared to the other agent designs in this problem variant.
%		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
%		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
%		?? REWORK/MAKE ACCURATE
		\label{fig:udp-tree-box}
	}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-single}

	\caption{
		Online performance for Opus benign traffic in a single-destination network, single-agent mode.
		?? DO I need this?
		\label{fig:udp-tree-box-single}
	}
\end{figure}

Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
\Cref{fig:udp-box} shows the behaviour for all choices of $n$---specifically, we see that Marl++ typically remains ahead of Marl, though this performance difference shrinks as $n$ increases, while SPF performs better at high host-density.

?? Discuss multi-dest topol once results available.

\subsection{Congestion-aware traffic}
%\begin{figure}
%	\centering
%	\includegraphics[width=0.9\linewidth]{../plots/tcp-2}
%	
%	\caption{
%		Online performance for $n=2$ hosts per egress point when benign traffic is TCP-like.
%		Marl++ and Marl achieve very similar performance, starting off similarly well without notable improvement over an episode.
%		SPF's performance is disappointingly close to baseline, indicating that it is as useful as having no defence system.
%		\label{fig:tcp-2}
%	}
%\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-separate}

	\caption{
		Online performance for HTTP benign traffic in a single-destination network, multi-agent mode.
		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for large $n$.
		\label{fig:tcp-tree-box}
	}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}

	\caption{
		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
		?? DO I need this?
		\label{fig-tcp-tree-box-single}
	}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}

	\caption{
		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
		At this level of host density, Guarded reaches higher peak performance sooner and is considerably more consistent throughout the episode.
		Guarded benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
		The Instant model converges to perform similarly to Marl.
%		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
		\label{fig:tcp-tree-16}
	}
\end{figure}

From \cref{fig:tcp-box}, Marl++ consistently allows more legitimate traffic to reach its destination than Marl, with both significantly above the baseline (\cref{tab:av-vals}, \emph{Capped}, TCP).
SPF performs slightly worse at low $n$, but outperforms both models as host density increases.
\Cref{fig:tcp-16} shows a trend in our results: for larger $n$, Marl++ requires around twice as long as Marl to learn its best-achieving policy, under-performing for this period, but achieves notably better peak and average performance.
SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.

?? Discuss multi-dest topol once results available.

\subsection{Single-agent performance}
Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---especially for TCP, and as $n$ grows larger (\cref{fig:tcp-16}).
Single-agent SPF almost consistently outperforms Marl.
We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.

@@

@@

%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
\fakepara{Model performance}
Of the results presented, SPF's unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
Accordingly, we see that SPF performs best when agents learn from as much knowledge as possible: high $n$ and single-agent training.
Why does it exhibit considerably lower variance?
Our belief is that SPF is, in fact, a considerably stronger model for these reasons, and its successes give us particular cause to consider the best schemes for efficient information sharing.
To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
Paradoxically, Marl++ generally achieves the best performance yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.

%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for.
%?? I don’t know if you want to also mention the problem of things getting “stuck” in locally optimal (but globally sub-optimal) policies. I think it’s fairly uncontroversial that an increase in state space would make this more likely. Also, the lack of discounting could play a large role (and more so as the state space becomes more complex…
The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
Our pre-training results support the idea that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.

%It is likely that the design of Marl++

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.

@@

%?? Is the state space interpretable? Yes!
Machine learning algorithms have earned a reputation for eluding human interpretation, while being vulnerable to evasion and poisoning.
Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
These signals' topological separation, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.

@@

Where overheads are concerned, the state space sizes guarantee that a Marl policy remains under \SI{427}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: \SIrange{16.5}{17.0}{\kibi\byte} from our experiments.
As we have described earlier, action updates require a constant number of floating point operations---\num{122} floating point additions and \num{2} multiplications per update of $\wvec{}$, above the \num{140} additions required to choose an action.
The vast majority of these operations can be vectorised trivially, if such hardware is present.
Action computation for SPF is cheaper still, requiring only \num{42} additions per action required.
Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
We believe that it may be possible to implement similar behaviour on standard commodity switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.

@@

?? Actually talk about Marl (???) to appease reviewer \#1.
The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15} (whom we have positioned our work against), and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
?? Recap their flaws, since they've been cut form every other aspect.

\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ISP networks, although some of our feature choices are backed by similar observations.
?? Flaws? what can't it do that we can do better? Long times to compute TBE routes on real networks, low expansion factor in real network can require more ``rounds'' of filtering. (``takes only'' \SI{14}{\second} for Cogent network? gross.)
?? I believe their assumptions may not extend to CBR traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic. Also DASH traffic.

@@

For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
We feel there is a shortage of work in this category at present, at least in the field of networks.
To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}, which we have covered extensively in \cref{sec:environment-and-rl-algorithm}.

@@

\section{Conclusions and Future Work}

Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.

>>>>

\begin{abstract}
%?? Trim to 75--200 words, simplify.
DDoS attacks plague the availability of online services today, yet like many cybersecurity problems are evolving and non-stationary.
Normal and attack patterns shift as new protocols and applications are introduced, further compounded by burstiness and seasonal variation.
Accordingly, application of machine learning-based techniques is difficult.
%Network intrusion detection and prevention systems backed by machine learning (and the autonomous operation they promise) have been long-awaited, but face problems hampering effective deployment.

\emph{Reinforcement learning} (RL) may overcome this detection problem for DDoS attacks by managing and monitoring \emph{consequences}; an agent's role is to learn to optimise performance criteria (which are always available) in an online manner.
%We examine how the paradigm has been applied to the task of \emph{Distributed Denial of Service} (DDoS) attack prevention to date, and present an analysis of suspected edge cases to uncover existing weaknesses in present approaches.
We advance the state-of-the-art in RL-based DDoS mitigation by introducing two agent classes designed to act on a per-flow basis, in a protocol-agnostic manner for any network topology.
%These flaws---poor handling of TCP traffic and high host density at egress points---inspire and direct an exploration of the design space of RL models.
This is supported by an in-depth investigation of feature suitability and empirical evaluation.

Our results show the existence of flow features with high predictive power for different traffic classes, when a basis for feedback-loop-like control.
We show that the new RL agent models can offer a significant increase in goodput of legitimate TCP traffic for many choices of host density.

@@

%?? Discuss the most important conclusions before the outline.
We find that the existing work for this task fails to account for congestion-aware traffic (i.e., TCP) and deployment environments with high host density per egress point, achieving poor results due to an overly coarse view of the network.
The approach we identify to remedy this is to make throttling decisions on a per-source basis, and we present the engineering decisions this mandates: updating RL agents from multiple traces per timestep, timed random sequential action computation and a supporting \emph{software-defined network} (SDN) architecture.
In tandem with the development and evaluation of an effective state-space and model, we provide the design of a second model inspired by past work on algorithmic DDoS prevention, as an example of the integration of domain-specific knowledge.
Our introduction of per-source decisions improves substantially upon the state-of-the-art when acting upon most internet traffic (i.e., congestion-aware protocols), and we show that our second model achieves excellent performance for high host density in this case.
Crucially, both models remain protocol- and content-agnostic to offer future-proofing against the rollout of future protocols like QUIC \cite{DBLP:conf/sigcomm/LangleyRWVKZYKS17}.
%?? Also algorithmic enhancements such as multiple actions per timestep, 
%?? PROTOCOL-AGNOSTIC -- HOW WILL THESE THINGS COPE WITH QUIC ET AL.?!?!
Briefly, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
%	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention (\emph{Instant} and \emph{Guarded} action models), improving upon past aggregate-based models (\cref{sec:ddos-mitigation-with-per-flow-reinforcement-learning}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Reactive simulations of HTTP and VoIP web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An empirical evaluation of our new models alongside the state-of-the-art in RL-based DDoS mitigation with these new traffic models (\cref{sec:the-results-of-doing-so}), alongside a discussion of security concerns and real-world deployment (\cref{sec:discussion}).
\end{itemize}

@@

\section{DDoS Mitigation with Per-flow Reinforcement Learning}\label{sec:ddos-mitigation-with-per-flow-reinforcement-learning}
Our main hypothesis is that the best method for advancing past the current shortcomings of RL-based DDoS mitigation is to modify the design of agents such that filtering decisions are computed per-flow.
However, any of these alterations must account for computational constraints imposed by the deployment environment---the amount of flows passing over an agent is unbounded.
We describe and justify our approach, our algorithmic improvements, and present two action models, one of which draws on domain knowledge introduced by SPIFFY \cite{DBLP:conf/ndss/KangGS16}.

@@

	\caption{
		Online performance for Opus benign traffic in a single-destination network, multi-agent mode.
%		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
		\emph{Instant} outperforms Marl for $n \in \{4, 8\}$ (with higher variance), but performs similarly to Marl at $n\in \{2, 16\}$.
		\emph{Guarded} always underperforms compared to the other agent designs in this problem variant.
%		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
%		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
%		?? REWORK/MAKE ACCURATE
		\label{fig:udp-tree-box}
	}
\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-udp-box-single}
%	
%	\caption{
%		Online performance for Opus benign traffic in a single-destination network, single-agent mode.
%		?? DO I need this?
%		\label{fig:udp-tree-box-single}
%	}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-udp-box-separate}
%	
%	\caption{
%		Online performance for Opus benign traffic in a multi-destination network, multi-agent mode.
%		?? DO I need this?
%		\label{fig:udp-ecmp-box}
%	}
%\end{figure}
%
%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-udp-box-single}
%	
%	\caption{
%		Online performance for Opus benign traffic in a multi-destination network, single-agent mode.
%		?? DO I need this?
%		\label{fig:udp-ecmp-box-single}
%	}
%\end{figure}

In a single-destination network, we observe that Marl's performance degrades as $n$ increases.
Typically, our \emph{Instant} agent design achieves the best performance in multi-agent mode, reducing collateral damage above the current state-of-the-art, but sharply degrades at low $n$ when agents share experience.
We see a reversal of this trend for the \emph{Guarded} model, which improves as $n$ increases and in single-agent mode---when $n\ge4$, the single-agent variant offers consistent improvement.
%Across all choices of $n$, we see that Marl++ exhibits reduced collateral damage compared to Marl, with SPF starting poorly yet becoming more effective for larger $n$ (\cref{tab:av-vals}, \emph{Capped}, UDP).
\Cref{fig:udp-tree-box} shows the preserved traffic in multi-agent mode.

?? Discuss multi-dest topol once all results available.

\subsection{Congestion-aware traffic}
%\begin{figure}
%	\centering
%	\includegraphics[width=0.9\linewidth]{../plots/tcp-2}
%	
%	\caption{
%		Online performance for $n=2$ hosts per egress point when benign traffic is TCP-like.
%		Marl++ and Marl achieve very similar performance, starting off similarly well without notable improvement over an episode.
%		SPF's performance is disappointingly close to baseline, indicating that it is as useful as having no defence system.
%		\label{fig:tcp-2}
%	}
%\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}

	\caption{
		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
		\emph{Instant} and \emph{Guarded} exhibit essentially identical efficacy at $n~=~2$, protecting less traffic than Marl.
		Only \emph{Guarded}'s performance rapidly increases with $n$, achieving considerably better median performance and lower variance than the other models.
		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
		\label{fig:tcp-tree-box}
	}
\end{figure}
%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-box-single}
%	
%	\caption{
%		Online performance for HTTP benign traffic in a single-destination network, single-agent mode.
%		?? DO I need this?
%		\label{fig-tcp-tree-box-single}
%	}
%\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.95\linewidth]{../plots/tnsm-tcp-16-single}

	\caption{
		Online performance of standard and single-agent models in a single-destination network with $n=16$ hosts per egress point and HTTP benign traffic.
		At this level of host density, \emph{Guarded} reaches higher peak performance sooner and is considerably more consistent throughout the episode.
		\emph{Guarded} benefits greatly from information sharing, converging to protect around \SI{75}{\percent} of TCP traffic within \SI{100}{\second}.
		The \emph{Instant} model converges Marl's level of performance.
%		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
		\label{fig:tcp-tree-16}
	}
\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-box-separate}
%	
%	\caption{
%		Online performance for HTTP benign traffic in a multi-destination network, multi-agent mode.
%		?? DO I need this?
%		\label{fig:tcp-ecmp-box}
%	}
%\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-box-single}
%	
%	\caption{
%		Online performance for HTTP benign traffic in a multi-destination network, single-agent mode.
%		?? DO I need this?
%		\label{fig:tcp-ecmp-box-single}
%	}
%\end{figure}

%\begin{figure}
%	\centering
%	\includegraphics[width=0.95\linewidth]{../plots/tnsm-ecmp-tcp-16-single}
%	
%	\caption{
%		?? Eh
%		\label{fig:tcp-ecmp-16}
%	}
%\end{figure}

\Cref{tab:av-vals} shows that Marl offers a low (though fairly consistent) level of protection for TCP traffic, which the \emph{Instant} agent offers no substantial improvement over.
However, \emph{Guarded} agents offer a remarkable improvement for this class of traffic, particularly when experience can be shared---offering a \SI{2.21}{$!\times$} improvement over the state-of-the art during training, which is made clearer in \cref{fig:tcp-tree-box}.
\Cref{fig:tcp-tree-16} shows that this model can protect a peak \SI{80}{\percent} of TCP traffic (\SI{2.5}{$!\times$} improvement) after just \SI{100}{\second}, but also that all of the new models require considerably longer than Marl to learn their best-achieving policy.
%SPF reaches a stronger plateau before \emph{both}, remaining more consistent and appearing to continue learning.

%?? Discuss multi-dest topol once results available.
We observe that the same trends present themselves in the multi-destination topology: \emph{Guarded} remains the best fit for TCP, in both training modes.
Crucially, the rigid tree of learners and teams which define Marl, along with its lack of action granularity, seem to be a poor fit in this environment.

%\subsection{Single-agent performance}
%Making all decisions with a single agent is roughly equivalent to having a zero-cost communication channel between each pair of agents, theoretically allowing faster training by giving each agent more experience.
%Curiously, we observe that this often leads to drastically worse policies when used as part of Marl++ for small $n$, but makes SPF a considerably more competitive model---especially for TCP, and as $n$ grows larger (\cref{fig:tcp-16}).
%Single-agent SPF almost consistently outperforms Marl.
%We discuss our conjectures for why this reversal occurs in \cref{sec:discussion}.

@@

%?? New angle -- \emph{Guarded} offers substantal improvements for \emph{most} internet traffic according to the stats we know. Is accurate protection of beningn UDP traffic actually more difficult?

@@

%?? Why does SPF only do well sometimes? Model is actually more difficult to learn, so it seems to do best when it has a larger set of decisions to learn from. But, it does worse for TCP?
\fakepara{Model performance}
Of the results presented, \emph{Guarded}'s unpredictable (often worse) starting performance is unexpected, given its far smaller action space.
It's natural to expect that this would make the model easier to learn, but the additional state required appears to make the task \emph{harder}, beyond even the value of choosing a non-zero discount factor (adding forward-planning to explicitly mitigate this effect).
Accordingly, we see that this design performs best (and exhibits considerably lower variance) when agents learn from as much knowledge as possible: high $n$ and single-agent training.
To filter incoming traffic from a source, it must decide to degrade their inbound traffic multiple times in a row, reducing the likelihood that a legitimate flow is punished by accident.
Our belief is that \emph{Guarded} is a considerably stronger model for these reasons, and its successes offer strong rationale to consider the best schemes for efficient information sharing.
Paradoxically, \emph{Instant} generally achieves the best performance for UDP traffic yet actively suffers when trained as a single learner---this may occur due to a roughly even spread of values between disparate actions, due to shared characteristics between legitimate and malicious flows.

%?? May be hard to learn multiple features at once while controlling multiple flows while contending with many more agents, with harder dynamics like TCP. Does this hinder learning in the long run?
Although we have improved upon Marl in both identified problem cases, the improvements are not quite on the order we'd expect, or hope for in the case of UDP traffic.
%?? I don’t know if you want to also mention the problem of things getting “stuck” in locally optimal (but globally sub-optimal) policies. I think it’s fairly uncontroversial that an increase in state space would make this more likely. Also, the lack of discounting could play a large role (and more so as the state space becomes more complex…
The most likely explanation is that agents are converging to, and becoming stuck in, locally optimal (but globally sub-optimal) policies.
The increased state-space size makes this a more likely occurrence, as is the unclear affect of hyperparameters ($\alpha$, $\gamma$) as we scale up the state-space.
We suspect that these difficulties may be exacerbated by the competitive nature of learning that these models embody: agents are learning action values for multiple features simultaneously, taking many actions at once (making it harder to observe the true value of each action), and controlling shared global state.
Moreover, benign UDP traffic shares many characteristics with attack traffic, suggesting that more training samples or some unknown feature might aid control, or that it may be worthwhile to extensively pre-train agents non-competitively on each feature using individual flows.

%It is likely that the design of Marl++ 
%?? Need to mention that Marl performs lower than their paper's numbers for TCP...

%Finally, it is crucial to note that the models and techniques presented here are an improvement , this work still trails behind existing (exact) DDoS flow detection mechanisms.
Most importantly, what we wish to impart is the knowledge that while the models and techniques we present here are a significant improvement over past RL-based work, this work still trails behind existing (exact) DDoS flow detection mechanisms.
Although we have conducted work to map the territory to some extent, there are still more advancements to be made before RL-based DDoS defence is truly competitive.
The benefits we have at present are, however, substantial.
What we offer above the alternate approaches we discuss in \cref{sec:related-work} are potentially more flexible deployments and low-overhead decision-making, without requiring active measurement or the network resources and capabilities that the most effective techniques rely upon.
The techniques we propose offer substantial improvements for the defence of congestion-aware traffic, which remains the most prevalent class of traffic on the internet today.
Moreover, our decision making processes are entirely agnostic of the protocol or content of inbound traffic, offering future-proofing against the introduction of new transports.

@@

%?? Is the state space interpretable? Yes!
Machine learning algorithms have earned a reputation for eluding human interpretation, while being vulnerable to evasion and poisoning.
Given the risks associated with introducing such techniques in the context of security, it is natural to be concerned with the interpretability of the models we have proposed.
With the exception of global state, the tile-coding parameters we make use of ensure that the set of all outputs for each new feature we add is relatively enumerable: for $n$ tilings and $c$ tiles per dimension there are $nc^{\dim{f}}$ individual action value vectors per feature $f$ (\num{48} for the new features we introduce, \num{10368} for global state), though considerably more combinations thereof ($c^{n \cdot \dim{f}}$).
%\Cref{fig:intern-16-tcp-iat,fig:intern-16-tcp-something} show how we may visualise the portion of a policy for each feature, and describe what information can be gained from doing so.
Furthermore, system state which is dependent on many signals drawn from across a wide network (such as our global state) is difficult to exert precise control over.
These signals' topological separation, in concert with their burstiness and unpredictability, may have substantial effects on an attacker's capabilities.

@@

Where overheads are concerned, the state space sizes guarantee that an \emph{Instant} agent's policy remains under \SI{520}{\kibi\byte}, although in practice our sparse representation typically leads to far smaller policies: $\sim$\SI{17.8}{\kibi\byte} from our experiments.
\emph{Guarded} policies are $3/10$ of the size.
As we have described earlier, action updates require a constant number of floating point operations---\num{160} floating point additions and \num{32} multiplications per update of $\wvec{}$ with per-tile updates, above the \num{160} additions required to choose an action.
The vast majority of these operations can be vectorised trivially, if such hardware is present.
Action computation for \emph{Guarded} agents is cheaper still, requiring only \num{48} additions per action required.
Beyond this, we require that egress switches are capable of co-hosting an agent (i.e., through \emph{network function virtualisation}), with the necessary hardware to support this choice.
We believe that it may be possible to implement similar behaviour on standard commodity switches through application of \emph{programmable data planes} \cite{DBLP:conf/ancs/JouetP17}.

@@

%?? Actually talk about Marl (???) to appease reviewer \#1.
The closest available approach within this field is that of \textcite{DBLP:journals/eaai/MalialisK15} (whom we have positioned our work against), and their contribution in applying RL to the task of intrusion prevention is significant: their work helps to show the viability of live, adaptive, feedback-loop-like control of the network to detect and prevent DDoS attacks.
They create a tree overlay topology (subdivided into teams), where each agent applies packet drop to \emph{all} flows inbound to a protected server.
%?? Recap their flaws, since they've been cut form every other aspect.
Our results show that their technique underperforms at high host density and when congestion-aware traffic dominates---that their results do not demonstrate this suggests an evaluation driven purely by traces (rather than live application dynamics).

\emph{SPIFFY} \cite{DBLP:conf/ndss/KangGS16} aims to remedy transit-link attacks by observing how flows from each source respond to a sudden increase in available bandwidth.
\Citeauthor{DBLP:conf/ndss/KangGS16} realise that bots participating in an attack are often unable to match this bandwidth expansion due to having already saturated the capacity of their outbound links, while legitimate flows typically speed up to match the new fair-share rate.
%Attackers must either be detected or reduce the throughput of each bot, increasing the cost of launching an attack.
Unlike our approach (and due to the class of attacks it is designed to defend against), SPIFFY is intended to be deployed within ASes, although some of our feature choices are backed by similar observations.
A weakness of their approach is that computing a route to measure bandwidth expansion on real networks can be costly (up to \SI{14}{\second} for the Cogent topology), and that the low expansion factors in real network can require more ``rounds'' of filtering. 
Their assumptions about traffic response to such bandwidth expansion may not extend to HTTP DASH traffic or constant bitrate traffic (e.g., UDP VoIP traffic), which we know makes up a sizeable proportion of network traffic.

@@

For categorising how RL fits into solving problems, we label works as direct- or indirect-control RL.
A \emph{direct-control} RL problem is one where the RL agent(s) are learning optimal control over a set of actions to act as the \emph{primary} defence or decision-maker---requiring measurements, reward functions and action sets tailored for this purpose.
%We feel there is a shortage of work in this category at present, at least in the field of networks.
To date, the best-fitting example we have encountered is that of \textcite{DBLP:journals/eaai/MalialisK15}.

@@

\section{Conclusions and Future Work}
Through this paper, we have discussed the paradigm of reinforcement learning and its relevance to network intrusion prevention.
Our belief is that the potential to learn feedback-loop-like control online and against non-stationary problems makes it particularly suited to the problems endemic to the field.
